---
Peer Assessment 1
---

LOADING AND PREPROCESSING THE DATA

Here we are going load the data.

```{r}
setwd("C:/Users/Sara/Desktop/Coursera/ReproducibleResearch/CourseProject/repdata-data-activity")
dat<-read.csv("activity.csv",header = TRUE,sep = ",", stringsAsFactors=FALSE, dec=".")
```

MEAN TOTAL NUMBER OF STEPS TAKEN PER DAY

1. Steps taken per day
```{r}
datsinna<-na.omit(dat)
library(plyr)
nsteps<- ddply(datsinna, .(date),summarise,Totalsteps = sum(steps, na.rm = TRUE))

```

2.If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day

The main difference between bar charts and histograms is that With bar charts, each column represents a group defined by a categorical variable; and with histograms, each column represents a group defined by a quantitative variable.
One implication of this distinction: it is always appropriate to talk about the skewness of a histogram; that is, the tendency of the observations to fall more on the low end or the high end of the X axis.
With bar charts, however, the X axis does not have a low end or a high end; because the labels on the X axis are categorical - not quantitative. As a result, it is less appropriate to comment on the skewness of a bar chart.

```{r}
library(ggplot2)
ggp <- ggplot(nsteps, aes(x=factor(nsteps$date), y=nsteps$Totalsteps, fill=city)) +geom_bar(aes(fill=nsteps$date),stat="identity") +labs(x="date", y=expression("total steps")) + labs(title=expression("Total steps per date"))
print(ggp)
```
3. Calculate and report the mean and median of the total number of steps taken per day

```{r}
meansteps<- ddply(datsinna, .(date),summarise,Totalsteps = mean(steps, na.rm = TRUE))
mediansteps<- ddply(datsinna, .(date),summarise,Totalsteps = median(steps, na.rm = TRUE))
```

AVERAGE DAILY ACTIVITY PATTERN

1.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
steps.interval <- aggregate(steps ~ interval, data=datsinna, FUN=mean)
library(lattice)
xyplot(steps ~ interval, steps.interval, type = "l",
    xlab = "Interval", ylab = "Number of steps")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
steps.interval$interval[which.max(steps.interval$steps)]
```

IMPUTING MISSING VALUES

1.Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
nmissingvalue<-sum(is.na(dat))
```
2.Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

3.Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
dat <- merge(dat, steps.interval, by="interval", suffixes=c("",".y"))
nas <- is.na(dat$steps)
dat$steps[nas] <- dat$steps.y[nas]
dat <- dat[,c(1:3)]
```

4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

The values differ from the first part of the assingent. Taking account missing data there is more complete information about the total daily number of steps

Histogram

```{r}
n2steps<- ddply(dat, .(date),summarise,Totalsteps = sum(steps, na.rm = TRUE))
ggp <- ggplot(n2steps, aes(x=factor(n2steps$date), y=n2steps$Totalsteps, fill=city)) +geom_bar(aes(fill=n2steps$date),stat="identity") +labs(x="date", y=expression("total steps")) + labs(title=expression("Total steps per date without missing values"))
print(ggp)
```

Mean and Median

```{r}
mean2steps<- ddply(dat, .(date),summarise,Totalsteps = mean(steps, na.rm = TRUE))
median2steps<- ddply(dat, .(date),summarise,Totalsteps = median(steps, na.rm = TRUE))
```

DIFFERENCES BETWEEN WEEKENDS AND WEEKDAYS

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
day <- weekdays(as.Date(dat$date))
daylevel <- vector()
for (i in 1:nrow(dat)) {
    if (day[i] == "sábado") {
        daylevel[i] <- "Weekend"
    } else if (day[i] == "domingo") {
        daylevel[i] <- "Weekend"
    } else {
        daylevel[i] <- "Weekday"
    }
}
dat$daylevel <- daylevel
dat$daylevel <- factor(dat$daylevel)
```
2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r}
steps.interval.day <- aggregate(steps ~ interval+daylevel, data=dat, FUN=mean)
library(lattice)
xyplot(steps ~ interval | daylevel, steps.interval.day, type = "l", layout = c(1, 2), xlab = "Interval", ylab = "Number of steps per kind of day")
```
